---
title: "Lab 5"
author: "Becca Reynolds"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives:

- Getting counts for different groups
- Use {lubridate} to parse dates
- Find confidence intervals & do t-tests with t.test()
- Heatmap with geom_tile()

```{r, include = FALSE}
# Attach packages

library(tidyverse)
library(here)
library(janitor)

```

```{r}
# Import Lobster Data

lobster_abundance <-
  read_csv(here::here("data", "lobster_abundance.csv"), 
                              na = "-99999") %>%
  clean_names()

```

Use dplyr::uncount() function to expand a frequency variable 

```{r}
lobster_tidy <- lobster_abundance %>%
  tidyr::uncount(lobster_count)

```

Now each lobster has its own row. This is tidy format. 


Let's look at this data a bunch of ways...aka exploratory data visualization
```{r}
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color = site),
              width = 0.2,
              alpha = 0.3)


```
Some of these have close to the same central tendency, CARP and IVEE have more outliers...

```{r}
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free") #facet wrap seperates graphs instead of having them all overlap each other on the same coordinate plane, scales = free will adjust the axes scales for comparison based on visualization

```
Now we want to plot actual observation values against perfectly standard normal distrubtions to see how our data compares

```{r}
ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color = site),
              alpha = 0.3)

ggplot(data = lobster_tidy, aes(sample = size_mm)) +
  geom_qq() +
  facet_wrap(~site, scales = "free")
```
We observe relativley linear results for AQUE. CARP is also rel. normally distributed but just a few outliers are there making it look strange on the high end. Same with IVEE...just some outliers, Vs. what Mohawk looked like in the histogram we surprisingly see a pretty linear outcome here and we can say with more confidence theres a normal distribution. NAPL is also displaying a normal distribution. So overal, the data is pretty normally distributed.


Let's lubridate to parse dates and times

add a new column using mutate() that contains my date as an actual date
```{r}
class(lobster_tidy$date)
#R sees the dates as characters, we want to change this. use mdy because that's the order in which the dates are currently stored
lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )
class(lobster_date$date_new)
```

Now we'll parse the date column to get the different pieces like month and year seperated

```{r}
lobster_parse_date <- lobster_date %>%
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),
    obs_year = lubridate::year(date_new)) # which makes a column containing the date as a number, but what if we want it in text? add label = TRUE
  
```

Count lobsters by different groupings...

?count()

I want to count the number of lobsters by year and month

```{r}

lobster_ym <- lobster_parse_date %>%
  count(obs_year, obs_month)

lobster_ym

lobster_ysite <- lobster_parse_date %>%
  count(obs_year, site)

lobster_ysite

lobster_site <- lobster_parse_date %>%
  count(site)
  
lobster_site

#group_by + summarize + n()
lobster_summary <- lobster_parse_date %>%
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_n = n())

lobster_summary


```


Let's look at some confidence intervals and t-tests

use t.test() function to find confidence intervals and perform t-tests

```{r}
ivee_lobsters <- lobster_tidy %>%
  filter(site == "IVEE") %>%
  pull(size_mm)

t.test(ivee_lobsters)

# What we just did is pretty much the safe way of doing the following, then we performed a t-test on it
my_vector <- c(3, 5, 9, 15, 24)
  

```


two sample t-test:
is there a significnt difference in mean lobster ranks between naples reef and mohawk reef?
```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>% 
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)
mohk_napl_ttest

```

?t.test

#scroll down in ?t.test help doc and look at value for more info

Mean Lobster size differed significantly between mohawk and naples reefs (t(`r mohk_napl_ttest$parameter`) = 19.849)...

Maybe an easier way...
```{r}
lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)

```



Make a geom_tile heatmap using lobster_ysite df
```{r}

ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))

```

